---
title: "Plot4"
author: "Jerez"
date: "11/29/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'asis')
```

### Course: Exploratory Data Analysis

Fine particulate matter (PM2.5) is an ambient air pollutant for which there is strong evidence that it is harmful to human health. In the United States, the Environmental Protection Agency (EPA) is tasked with setting national ambient air quality standards for fine PM and for tracking the emissions of this pollutant into the atmosphere. Approximatly every 3 years, the EPA releases its database on emissions of PM2.5. This database is known as the National Emissions Inventory (NEI). You can read more information about the NEI at the [EPA National Emissions Inventory web site ](https://www3.epa.gov/ttn/chief/eiinformation.html)

For each year and for each type of PM source, the NEI records how many tons of PM2.5 were emitted from that source over the course of the entire year. The data that you will use for this assignment are for 1999, 2002, 2005, and 2008.

## Library

```{r}
setwd("~/Google Drive/Coursera/EDA")

# Install and load required packages
if (!require('ggplot2'))      {install.packages('ggplot2')}
if (!require('dplyr'))      {install.packages('dplyr')}
if (!require('data.table')) {install.packages('data.table')}

library(ggplot2)
library(dplyr)
library(data.table)

````


## Download the dataset

```{r}
# name for zip file
setwd("~/Google Drive/Coursera/EDA")
file.zip <- 'EDA_Final.zip' 

# Cheking if zip file exists
if (!file.exists(file.zip)){
 file.URL <- 'https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip'
 download.file(file.URL, file.zip, method = 'curl')
 unzip(file.zip, exdir = '.')} 

NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")

`````

# Question 4

Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

```{r}

SCC.coal = SCC[(grepl(x = SCC$Short.Name, pattern = "Coal", ignore.case=TRUE)),]
NEI.coal = merge(NEI, SCC.coal, by = "SCC")

Emissions <- NEI.coal %>% 
 group_by(year) %>%
 summarize(Total = sum(Emissions, na.rm = TRUE)) # Removing the NA's

Emissions$group = rep("coal",nrow(Emissions))

````
```{r}
Plot4 <- ggplot(Emissions, aes(as.factor(year), Total, group = group, label = round(Total,digits = 0)))+
 geom_point(size = 5, alpha = 0.6, color = 'violet')+
 geom_text(check_overlap = TRUE, family = 'Times New Roman', hjust = 1.3)+
 labs(x = 'Years', y = 'Total emissions (tons)', 
      title = 'Total Annual Coal combustion emissions in the US from 1998 to 2008')+
 theme_bw()

Plot4

ggsave('Plot4.png', plot = Plot4, width = 18, height = 9, units = 'cm')

````



## Create Plot4.R


````{r}
#library(knitr)
#purl('Plot4.Rmd')
````


